### 2017-03-16 R/tutorial.Rmd
Update to tutorial

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```
To begin using ITHIM one first must install it.  The package devtools
makes this is easy.

@@

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
#install_github("syounkin/ITHIM", ref="master")
install_github("syounkin/ITHIM", ref="devel")
library("ITHIM")
```

@@

We begin by creating an ITHIM object to represent the baseline model.
Currently ITHIM reads three files for the necessary input parameters.
Sample files are listed below.

@@

These files can be found in the master branch of ITHIM repository.

@@

Using the list of three file names we can create the ITHIM object using `createITHIM`.

#### Active Transport File Format
```
ageClass,sex,mode,value
ageClass1,M,walk,10
ageClass2,M,walk,9
...
ageClass7,F,cycle,15
ageClass8,F,cycle,13
```
The rows must be ordered by ageClass.
#### GBD File Format
The GBD file format has not been normalized like the others yet.
```
disease,sex,ageClass,dproj,yll,yld,daly
BreastCancer,M,ageClass1,0,0,0,0
BreastCancer,M,ageClass2,0,0,0,0
...
RTIs,F,ageClass7,1486.647682,21930.32348,7921.96966,29852.3031
RTIs,F,ageClass8,1957.865757,10720.47109,5222.446302,15942.95826
```
RTIs stands for road traffic injuries.

@@

### Road Injury Data

@@

#### Road Injury File Format
* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)

#### Road Injuries File Format
```
severity,roadType,strikingMode,victimMode,value
Fatal,Local,walk,walk,0
Fatal,Local,cycle,walk,13
...
Serious,Highway,LGV,NOV,0
Serious,Highway,mbike,NOV,0
```
ebikes are not included

@@

>>>>

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```
To begin using ITHIM one first must install it.  The package
_devtools_ makes this is easy.  Install _devtools_ using
`install.packages("devtools")` or some other way in _RStudio_.

@@

We use the function `install_github` to install the _master_ branch of
the _ITHIM_ package repository.

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("syounkin/ITHIM", ref="master")
library("ITHIM")
```

@@

We begin by creating an ITHIM object to represent the baseline model.
Currently _ITHIM_ can read up to three files containing input
parameters, but it is not necessary to provide all three.  Two sample
files are listed below that contain model parameters for the active
transport component of _ITHIM_.

@@

For a description of the file format see the help page for
`createITHIM` using `help("createITHIM")`.

@@

### File Formats
#### Active Transport File Format

The active transport file contains mean walking and cycling times
stratified by sex and _ITHIM_ age class.  The default file is located
in _inst/_.

```
ageClass,sex,mode,value
ageClass1,M,walk,21.47
ageClass2,M,walk,53.11
ageClass3,M,walk,50.13
...
ageClass6,F,cycle,1.53
ageClass7,F,cycle,0.53
ageClass8,F,cycle,0.23
```

Notes about file format:
* The rows must be ordered by ageClass.
* The mean values are in minutes per week

#### Disease Burden File Format

The disease burden file contains estimates for disease burden
stratified by _ITHIM_ age class and sex.

```
disease,sex,ageClass,burdenType,value
BreastCancer,M,ageClass1,dproj,0
BreastCancer,M,ageClass2,dproj,0
BreastCancer,M,ageClass3,dproj,0
...
RTIs,F,ageClass6,daly,49108.81151
RTIs,F,ageClass7,daly,29852.3031
RTIs,F,ageClass8,daly,15942.95826
```

Notes about file format:
* RTIs stands for road traffic injuries
* Only _ITHIM_ diseases can be included, namely "BreastCancer",
  "ColonCancer", "CVD", "Dementia", "Diabetes", "Depression" and "RTIs"

@@

#### Road Injuries File Format

@@

* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/master/inst/roadInjuries.csv)

```
severity,roadType,strikingMode,victimMode,value
Fatal,Local,walk,walk,0
Fatal,Local,cycle,walk,13
Fatal,Local,bus,walk,213
...
Serious,Highway,HGV,NOV,0
Serious,Highway,LGV,NOV,0
Serious,Highway,mbike,NOV,0
```

Notes about file format:
* ebikes are not included

@@

## Built-In Example

An example is contained within the _ITHIM_ package.

```{r misc225, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
example("ITHIM")
```
